\blue{...Numerical investigation of viscous flow instabilities through heterogeneous porous media using adaptive and unstructured meshes under different mobility scenarios...}

### 2017-07-04 Papers/ViscousInstabilities/Manuscript_ViscousInstabilities_KC.tex
updates

\medskip
Viscous finger morphologies were investigated by \citet{guan_2003} based on mathematical formulation and semi-analytic solutions of the coupled Darcy and continuity equations developed by \citet{mclean_1981}. A sensitivity study was conducted to assess the impact on fingers' formation, dimensions and branchiness for a range of viscosity ratio $\left(\right.$10$^{2}\le\textit{VR}\le$10$\left.^{4}\right)$ and modified capillary number $\left(\right.$632$\le N_{C}^{'}\le$6.32$\times$10$^{7}$, with $N_{C}^{'}= U_{f}\mu\gamma^{-1}(W/b)^{2}$, where $U_{f}$ is the velocity of the finger, $W$ is half Hele-Shaw cell width and $b$ is the cell thickness$\left.\right)$. Finger width $\left(\lambda_{f}\right)$ and volumetric flow rate $\left(\mathcal{Q}\right)$ at the outflow region are correlated through \citep{guan_2003},    
\begin{equation}   
   \mathcal{Q} = U_{f} b \lambda_{f}.\label{eqn:guaneqn} 
\end{equation}  
Here, numerical simulations performed with viscosity ratio of 10 and 150 indicated fingers width of approximately 0.45-0.70 and 0.50-0.90 cm (Fig.~\ref{fig:homoheleshaw_VN10_VN150} a and b), respectively. This is in close agreement with expected values obtained from Eqn.~\ref{eqn:guaneqn} which indicates finger width ranging from 0.13 to 0.75 (for {\it VR}=10) and from 0.30 to 0.75 (depending on finger's tip velocity, Table~\ref{table:Heleshaw}).

\medskip
Fluid flow dynamics through heterogeneous porous media are sensibly more complex than in homogeneous media and strongly depend on problem properties. Spatial variation in geological formations occurs in all length-scales, where characteristics of heterogeneities in small (pore) length-scales are statistically embedded into permeability (absolute and relative) and porosity parameters. Such multi-scale heterogeneity induces preferential flow pathways and plays a significant role in the onset of fluid instabilities as it triggers fingers formation and their accelerated growth \citep{ewing_1989,tchelepi_1994}.

In order to qualitatively investigate the impact of heterogeneity (here represented by changing in the permeability field), numerical simulations were conducted using a prescribed permeability distribution and viscosity ratios of 10 and 150. Absolute permeability ranging from 1.0$\times$10$^{-12}$ to 5.0$\times$10$^{-10}$ cm$^{2}$ (\ie 0.1 $\le$ {\bf K} $\le$ 50 milidarcy) was used in the simulations, which were performed with the same mesh resolution, boundary and initial conditions as in previous homogeneous cases. Figures~\ref{fig:HeleShawHeter_VR10} and ~\ref{fig:HeleShawHeter_VR150} show formation of multiscale elongations with continuous growth and coalescence of dendritic fingers in simulations performed with viscosity ratio of 10 and 150, respectively.

Finger width of approximately \red{XXX-XX} cm were obtained \red{(JG: COMPARE IT W THEORETICAL VALUE)}.

>>>>

\medskip
Viscous finger morphologies were investigated by \citet{guan_2003} based on mathematical formulation and semi-analytic solutions of the coupled Darcy and continuity equations developed by \citet{mclean_1981}. A sensitivity study was conducted to assess the impact on fingers' formation, dimensions and branchiness for a range of viscosity ratio $\left(\right.$10$^{2}\le\textit{VR}\le$10$\left.^{4}\right)$ and modified capillary number $\left(\right.$632$\le N_{C}^{'}\le$6.32$\times$10$^{7}$, with $N_{C}^{'}= U_{f}\mu\gamma^{-1}(W/b)^{2}$, where $U_{f}$ is the velocity of the finger, $W$ is half Hele-Shaw cell width and $b$ is the cell thickness$\left.\right)$. Finger width $\left(\lambda_{f}\right)$ and volumetric flow rate $\left(\mathcal{Q}\right)$ at the outflow region are correlated through \citep{guan_2003},    
\begin{equation}   
   \mathcal{Q} = U_{f} b \lambda_{f}.\label{eqn:guaneqn} 
\end{equation}  
Here, numerical simulations performed with viscosity ratio of 10 and 150 indicated maximum fingers width of approximately 0.45-0.70 and 0.50-0.90 cm (Fig.~\ref{fig:homoheleshaw_VN10_VN150} a and b), respectively. This is in close agreement with expected values obtained from Eqn.~\ref{eqn:guaneqn} which indicates maximum finger width ranging from 0.13 to 0.75 (for {\it VR}=10) and from 0.30 to 0.75 (depending on finger's tip velocity, Table~\ref{table:Heleshaw}).

\medskip
Fluid flow dynamics through heterogeneous porous media are sensibly more complex than in homogeneous media and strongly depend on problem properties. Spatial variation in geological formations occurs in all length-scales, where characteristics of heterogeneities in small (pore) length-scales are statistically embedded into permeability (absolute and relative) and porosity parameters. Such multi-scale heterogeneity induces preferential flow pathways and plays a significant role in the onset of fluid instabilities as it triggers fingers formation and their accelerated growth \citep[see][]{ewing_1989,tchelepi_1994}.

In order to qualitatively investigate the impact of heterogeneity (here represented by changing in the permeability field), numerical simulations were conducted using a prescribed permeability distribution and viscosity ratios of 10 and 150. Absolute permeability ranging from 1.0$\times$10$^{-12}$ to 5.0$\times$10$^{-10}$ cm$^{2}$ (\ie 0.1 $\le$ {\bf K} $\le$ 50 milidarcy) was used in the simulations, which were performed with the same mesh resolution, boundary and initial conditions as in previous homogeneous cases. Figures~\ref{fig:HeleShawHeter_VR10} and ~\ref{fig:HeleShawHeter_VR150} show formation of multi-scale elongations with continuous growth and coalescence of dendritic finger branching in simulations performed with viscosity ratio of 10 and 150, respectively. In both simulations, fingers width of approximately \red{XXX-XX} cm were obtained \red{(JG: COMPARE IT W THEORETICAL VALUE)}.

### 2017-07-04 Papers/ViscousInstabilities/article_table.tex
updates

%%%
%%%   TABLE 2
%%%
\begin{table}
  \begin{tabular}{c | c c c c}
    \hline\hline
                                  & {\bf K}                                  & {\it VR}  & $\lambda_{f}^{\text{(calc)}}$ (cm) & $\lambda_{f}^{\text{(simul)}}$ (cm) \\
                                  & $\left(\text{cm}^{2}\right)$              &           &  (Eqn.~\ref{eqn:guaneqn})  &                    \\
    \hline\hline
    {\bf Case 1} (homogeneous)    &  1$\times$10$^{-10}$                      &    10      &  0.13-0.75                 &   0.45-0.70       \\
    {\bf Case 2} (homogeneous)    &  1$\times$10$^{-10}$                      &    150     &  0.30-0.75                 &   0.50-0.90       \\
    {\bf Case 3} (heterogeneous)  &  1$\times$10$^{-12}$-5$\times$10$^{-10}$   &    10      &  0.25-4.50                 &   0.44-0.56       \\
    {\bf Case 4} (heterogeneous)  &  1$\times$10$^{-12}$-5$\times$10$^{-10}$   &    150     &  0.88-3.00                 &   0.24-0.40       
  \end{tabular}
   \caption{Flow simulations in Heleshaw cells (Section~\ref{section:results_homo_hete}): summary of finger width calculations using Eqn.~\ref{eqn:guaneqn} and obtained through numerical simualtions. $\lambda_{f}^{\text{calc}}$ was calculated based on fingers' tip velocities and outflow rate. Calculations were based on simulation data obtained from snapshots shown in Fig.~\ref{fig:homoheleshaw_VN10_VN150}.}\label{table:Heleshaw}
\end{table}

>>>>

%%%
%%%   TABLE 2
%%%
\begin{table}
  \begin{tabular}{c | c c c c}
    \hline\hline
                                  & {\bf K}                                  & {\it VR}  & $\lambda_{f}^{\text{(calc,max)}}$ (cm) & $\lambda_{f}^{\text{(simul,max)}}$ (cm) \\
                                  & $\left(\text{cm}^{2}\right)$              &           &  (Eqn.~\ref{eqn:guaneqn})  &                    \\
    \hline\hline
    {\bf Case 1} (homogeneous)    &  1$\times$10$^{-10}$                      &    10      &  0.13-0.75                 &   0.45-0.70       \\
    {\bf Case 2} (homogeneous)    &  1$\times$10$^{-10}$                      &    150     &  0.30-0.75                 &   0.50-0.90       \\
    {\bf Case 3} (heterogeneous)  &  1$\times$10$^{-12}$-5$\times$10$^{-10}$   &    10      &  0.25-4.50                 &   0.44-0.56       \\
    {\bf Case 4} (heterogeneous)  &  1$\times$10$^{-12}$-5$\times$10$^{-10}$   &    150     &  0.88-3.00                 &   0.24-0.40       
  \end{tabular}
   \caption{Flow simulations in Heleshaw cells (Section~\ref{section:results_homo_hete}): summary of finger width calculations using Eqn.~\ref{eqn:guaneqn} and obtained through numerical simualtions. $\lambda_{f}^{\text{calc}}$ was calculated based on fingers' tip velocities and outflow rate. Calculations were based on simulation data obtained from snapshots shown in Fig.~\ref{fig:homoheleshaw_VN10_VN150}.}\label{table:Heleshaw}
\end{table}
